Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-10 fib2.tex
Some minor wordsmithing.

We consider a 
2-dimensional system with
point-like excitations or quasiparticles known as anyons.
These anyons show exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the \emph{observables} of the system: these
are regions bounded by circles (Wilson loops)
and measure the total charge of the region.
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. % (UMTC).

@@

From the perspective of
\emph{topological quantum field theory}
(TQFT), anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid, but the space
around the anyons that becomes braided (twisted).
The generators corresponding to the braid
group generators are called half-twists.
This is the perspective we adopt.
In particular, this shows the action of braiding
on arbitrary observables (Wilson loops) as
these occupy the space around the anyons.

The state of a collection of anyons
can be described by the observables of the system.
An important difference between abelian and non-abelian
theories is this: the outcome of measurements in an abelian theory
are determined by particle content only.
That is, the fusion space is one-dimensional (a phase).
This is no longer the case in a non-abelian theory.
%An example is Fibonacci: $\tau\times\tau=\tau+\vac$
%\emph{picture here?}
%specify a state requires nesting multiple wilson lines.
In particular, we consider a system supporting Fibonacci
anyon excitations, denoted by $\tau.$
Two such anyons can have total charge
that is a superposition of $\tau$ and $\vac$ (vacuum),
so the fusion space in this case is 2-dimensional:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
the golden ratio. Additionally, there is a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We will consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
%For convenience 
%(in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), 
This global space is the one in which we will encode, and demonstrate protection of, quantum 
information.

Non-intersecting loops commute, and so
a basis for the state can be built from a maximal
set of disjoint, nested loops.
For three anyons, we relate two
possible ways of nesting loops %these measurements
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lhs shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The rhs involves a superposition (summation)
over charge states $j$ of the anyons with charge $b$ and $c$.
The green line serves to linearly
order the anyons, and track
deformations in the underlying space 
(the history of braiding processes).
It also denotes the direction along
which $F$-moves occur.

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways.
\dude{picture? ref Fig 2?}

Braiding within an observable will only change the phase
of that observable \dude{???}, otherwise may twist these
loops in various ways:
\begin{align*}
\includegraphics[]{pic-braid-loop.pdf}
\end{align*}

These pictures, abstracted from the quantum context,
are called \emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
In general, such a curve diagram is 
a continuous embedding of a line that passes through 
each puncture once. 
Remarkably, any two such linear orderings are related by
braiding (half-twists).
\dude{optional paragraph..}

The $R$ and $F$-moves obey composition of diagrams.
\dude{picture?}
Fusion is the ``reverse'' of composition: replace the
interior of a disc by a single charge.
\dude{picture?}

@@

%We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
%microscopic details of the system. 
%This is consistent with the principles of topologically 
%ordered systems and anyonic physics, where the key universal features describing the 
%anyon model correspond to large length-scale physics, while the microscopic physics plays 
%a less important (and non-universal) role.
%Our simulation is in effect that of a $(2+1)$-dimensional topological 
%quantum field theory, which allows for topological processes such as 
%pair-creation, braiding, and fusion of anyonic charges. 
%This is modeled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
%This treatment applies equally well to lattice models, where anyons 
%can only be located at a discrete set of points, 
%or continuum models, where a discretization of the manifold must 
%be chosen in order to perform charge measurements. 
%In the former case, each tile of $\Lambda$ represents a 
%lattice site, while in the latter case each tile represents 
%a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

@@

@@

At some point, with noise strong enough, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, this regime is somewhere
beyond the decoder threshold, which is the 
point at which the decoder creates nontrivial (extensive) loops.
We can view the system state as a tensor network, on which
the decoder is performing an (efficiently simulable) contraction.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
\dude{more polish needed...}

%Of course, if we were to consider noise so strong that 
%the components percolate over the entire lattice, we would no 
%longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
%However, recall that logical errors in our system correspond to 
%processes that act over a non-trivial loop of the torus. 
%Below the point at which the decoder begins to create 
%such nontrivial loops, our simulation remains efficient
%while it begins to declare failure whenever non-trivial logical 
%operations occur. This allows us to simulate large lattice sizes 
%in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

In order to keep track of the state of the system we maintain a set of disjoint curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
\dude{Heisenberg picture?}
The observable corresponding to the tile is then
represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}

>>>>

We consider a 
2-dimensional system with
point-like excitations or quasiparticles known as anyons.
These anyons show exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the topological observables of the system: these
are total charge measurements within regions bounded by closed loops (Wilson loops).
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. % (UMTC).

@@

From the perspective of
\emph{topological quantum field theory}
(TQFT), anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid, but the manifold in which the anyons embed 
is considered deformed (twisted) by these braid moves.
%The generators corresponding to the braid
%group generators are called half-twists.
% -we thought this was a getting ahead of ourselves since we haven't introduced this stuff yet.
This is the perspective we adopt.
In particular, this shows the action of braiding
on arbitrary observables (Wilson loops) as
these deform with the manifold.

The state of a collection of anyons
can be described by the observables of the system.
The difference between abelian and non-abelian
anyon theories is that the outcomes of measurements in an abelian theory
are determined uniquely particle content only.
That is, the fusion space is one-dimensional for abelian anyons, and generally larger for non-abelian anyons.
%An example is Fibonacci: $\tau\times\tau=\tau+\vac$
%\emph{picture here?}
%specify a state requires nesting multiple wilson lines.
In particular, we consider a system supporting (non-abelian) Fibonacci
anyon excitations, denoted by $\tau.$
Two such anyons can have total charge
that is a superposition of $\tau$ and $\vac$ (vacuum),
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams of definite total charge for the Wilson loops, and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}

For $n$ anyons of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
the golden ratio. Additionally, there is a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We will consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
%For convenience 
%(in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), 
This global space is the one in which we will encode, and demonstrate protection of, quantum 
information.

Observables associated to non-intersecting loops commute, and so
a basis for the space can be built from a maximal
set of disjoint, nested loops.
For three anyons, two
possible ways of nesting loops are related %these measurements
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lefthand side shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The righthand side involves a superposition 
over total charge $j$ of the anyons with charge $b$ and $c$.
The green line serves to linearly
order the anyons, and track
deformations in the underlying space 
(the history of braiding processes).
It also denotes the direction along
which $F$-moves occur.

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways.
\dude{picture? ref Fig 2?}
%
%
%Braiding within an observable will only change the phase
%of that observable \dude{???}, otherwise may twist these
%loops in various ways:
\begin{align*}
\includegraphics[]{pic-braid-loop.pdf}
\end{align*}
%
%
These diagrams can be composed, and the fusion operation replaces the 
interior of a disc by a single charge.
These pictures, abstracted from the quantum context,
are called \emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
%In general, such a curve diagram is 
%a continuous embedding of a line that passes through 
%each puncture once. 
%Remarkably, any two such linear orderings are related by
%braiding (half-twists).
%\dude{optional paragraph..}

@@

%We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
%microscopic details of the system. 
%This is consistent with the principles of topologically 
%ordered systems and anyonic physics, where the key universal features describing the 
%anyon model correspond to large length-scale physics, while the microscopic physics plays 
%a less important (and non-universal) role.
%Our simulation is in effect that of a $(2+1)$-dimensional topological 
%quantum field theory, which allows for topological processes such as 
%pair-creation, braiding, and fusion of anyonic charges.

@@

We model the system on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
This treatment applies equally well to lattice models, where anyons 
can only be located at a discrete set of points, 
or continuum models, where a discretization of the manifold must 
be chosen in order to perform charge measurements. 
In the former case, each tile of $\Lambda$ represents a 
lattice site, while in the latter case each tile represents 
a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

@@

At some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, this regime is somewhere
beyond the decoder threshold, which is the 
point at which the decoder creates nontrivial (extensive) loops.
\stf{CGB and I are still not confident in the previous sentence or the next one}
We can view the system state as a tensor network, on which
the decoder is performing an (efficiently simulable) contraction.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
\dude{more polish needed...}

%However, recall that logical errors in our system correspond to 
%processes that act over a non-trivial loop of the torus. 
%Below the point at which the decoder begins to create 
%such nontrivial loops, our simulation remains efficient
%while it begins to declare failure whenever non-trivial logical 
%operations occur. This allows us to simulate large lattice sizes 
%in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

In order to keep track of the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
\dude{Heisenberg picture?}
The observable corresponding to the tile is then
represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}
